---
title: Durational measurements on the Dutch prefix *ge-*
author: "Felix Golcher and Alter Ego"
date: "`r Sys.Date()`"
fontsize: 10pt
output:
  bookdown::pdf_document2:
    keep_tex:  true
    toc: false
  html_document: default
  bookdown::word_document2:
    toc: false
    fig_caption: yes
bibliography: bib.bib
---

```{r setup, include=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = F,
                      message = F,
                      cache = F,
                      autodep = T)
options(digits=3)

inline_hook <- function(x) {
  if (is.numeric(x)) {
    knitr:::format_sci(x, 'html')
  } else x
}
knitr::knit_hooks$set(inline = inline_hook)

```

```{r, include=F}
data("durationsGe", package = "languageR")
library(tidyverse)
```

```{r, include=FALSE, eval=T}
durationsGe <- durationsGe %>% filter(!is.na(Speaker))
```


```{r, include=FALSE, eval=T}
durationsGe <- durationsGe %>% filter(Speaker != "N01033")
```


**Abstract:** This is about Dutch prefixes. It uses an example data set from [@baayen08].

```{r, include=FALSE}
m0 <- lm(DurationOfPrefix~Frequency, durationsGe)
m0s <- summary(m0)
m0c <- confint(m0)
```

# Introduction

```{r blub, echo=FALSE, fig.cap="A very interesting graphical description of the data.", fig.width=5, fig.height=2.5}
durationsGe %>% as_tibble() %>% 
  ggplot(aes(Frequency, DurationOfPrefix))+
  geom_point()+
  scale_x_log10()+
  scale_y_log10()+
  geom_smooth(method=lm)+
  theme_bw()
```


We tried to figure out if the frequency of a word has an influence on the duration of Dutch prefixes. Intuitively one could assume that with the duration drops with frequency. For this we recruited `r length(unique(durationsGe$Speaker))` Dutch native speakers. They produced `r length(unique(durationsGe$Word))` different words.

Figure \@ref(fig:blub) gives an overview of the data. We clearly see a downward trajectory.

We actually made a significance test. The estimate for the slope is `r coefficients(m0s)[2,1]` with a confidence interval being [`r m0c[2,1]`, `r m0c[2,2]`]. This is not the most impressive result, adn puts the $p$-value of `r signif(coefficients(m0s)[2,4],3)` a bit into perspective.

# References