---
title: "Reuse Code"
subtitle: "A Small Presentation"
author: "You and Me"
institute: "The Invisible University"
date: "`r Sys.Date()` (today)"
output:
  xaringan::moon_reader:
    self_contained: true
    _lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r, include=FALSE}
library(ggplot2)
```


## Task description

* The last solution had a drawback:
  - The same code was repeated for the second version of the chunk
  - If you change the first version and forget the second, it'l be inconsistent.
* Reuse the code of the first block explicitely in the second.
  - Use the chunk option `ref.label` for that.
  - Remember: You have to give the first chunk a name for that.

---

## Reuse Case 1

.pull-left[
```{r goodfig, eval=F}
ggplot(diamonds,
       aes(carat, price, 
           col=color))+
  geom_point()
```
]

.pull-right[
```{r, ref.label="goodfig", echo=FALSE}
```
]

If everything went fine, every change in the left chunk pertains to the right as well.